---
title: "CalculatingES50"
author: "Natalie Posdaljian"
date: "8/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Hurlbert's index, or ES50, is the expected number of distinct species in a random sample of 50 observations; it is a bias-indepenent indicator on marine biodiversity richness. It has the ability to somewhat correct for sampling bias.

Assumptions of ES50
- That indivdiuals are randomly distributed
- Sample size is sufficiently large
- Samples are taxonomically similar
- All samples have been taken in the same manner

```{r}
#loading robis
installed <- rownames(installed.packages())
if ( !("robis" %in% installed) ){
   if ( !("remotes" %in% installed) )install.packages("remotes")
   remotes::install_github("iobis/robis")
}

#load libraries
library(robis)
library('ggplot2')
library("rnaturalearth")
library("rnaturalearthdata")
library(roperators)
library(dplyr)
library(dggridR)
library(magrittr) 
```

Load all occurences within a specificed polygon. (WARNING: using Casco Bay for now, but this will eventually be the MPA polygons)

```{r}
CascoBay <- occurrence(geometry = "POLYGON ((-70.2 43.5, -69.8 43.5, -69.8 43.9, -70.2 43.9, -70.2 43.5)) ")
```

User Specified metrics
```{r}
rank <- "scientificName"
```

