---
title: "obis_mpa_sel_country"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sf)
library(dplyr)
library(robis)
library(leaflet)
library(data.table)
source("R-code/wdpar-package.R")

# robis::occurrence() expect filtering polygons in a well-known-text represerntation (aka WKT)
extract_polygon_geometry <- function(polygon) {
  polygon %>%
    sf::st_geometry() %>%
    sf::st_as_text()
}

```

## Grabs global WDPA dataset and filters by selected country using ISO3 codes (https://unstats.un.org/unsd/tradekb/knowledgebase/country-code)

```{r}
country_mpa <- wdpa_read_global(filestream=TRUE) %>%
  dplyr::filter(ISO3 %in% "BEL")
```


## Extract occurence data from OBIS within MPAs for selected country (takes a long time for data rich MPAs!)

```{r obis}

# gets the occurrence data for a single multipolygon (one row of WDPA table)
# get_one_mpa <- function(x, key){
#    robis::occurrence(geometry = sf::st_as_text(sf::st_convex_hull(sf::st_geometry(x))))
# }
# 
# df <- country_mpa %>%
#   dplyr::rowwise() %>%
#   dplyr::group_map(get_one_mpa,
#                    .keep = TRUE) %>%
#   dplyr::bind_rows()
   
```

## Visualize occurrence records from OBIS for selected country (takes a long time for data rich MPAs!)

```{r}
# leaflet() %>%
#   addTiles("https://cartodb-basemaps-{s}.global.ssl.fastly.net/light_all/{z}/{x}/{y}.png") %>%
#   addCircleMarkers(lat = df$decimalLatitude, lng = df$decimalLongitude, radius = 5, weight = 0, fillOpacity = 1, fillColor = "#cc3300")

```

## Extracts OBIS records within all MPAs for selected country

```{r}
convert_mpa_to_WKT <- function(x){
  st_as_text(st_convex_hull(st_geometry(x)))
}

get_datasets <- function(x){
  get_one_dataset <- function(x, key){
    robis::dataset(geometry=convert_mpa_to_WKT(x))
  }
  x %>% 
    dplyr::rowwise() %>% 
    dplyr::group_map(get_one_dataset, .keep=TRUE) %>% 
    dplyr::bind_rows()
}

##obis_rec <- get_datasets(country_mpa)
```

## This function needs to be run before the next code chunk

```{r}
  get_one_dataset <- function(x, key){
    robis::dataset(geometry=convert_mpa_to_WKT(x))
  }
```

## Extracts OBIS records by looping through the list of MPAs for selected country

```{r}
rec_summ <- matrix(ncol = 2, nrow = nrow(country_mpa)) 

for (j in 1:nrow(country_mpa)){
  mpa_rec <- country_mpa$geom[j] %>% get_one_dataset()
    if(nrow(mpa_rec) == 0)
      num_rec <- 0
    if(nrow(mpa_rec) > 0)
      num_rec <- sum(mpa_rec$records)
      mpa_id <- country_mpa$WDPA_PID[j]
      rec_summ[j, ] <- cbind(mpa_id, as.integer(num_rec))
}
rec_summ <- data.frame(rec_summ) 
colnames(rec_summ) <- c("WDPAID", "records")
```

## This adds number of OBIS records to 'country_map' object

```{r}
country_mpa2 <- cbind(country_mpa, rec_summ$records) %>% setnames(old = c('rec_summ.records'), new = c('records'))
country_mpa2$records <- sapply(country_mpa2$records, as.numeric)
```

## Generates map with MPAs with WDPAID and number of OBIS records

```{r}
bins <- c(0, 1000, 10000, 100000, 200000, Inf)
pal <- colorBin("YlOrRd", domain = country_mpa2$records, bins = bins)

leafletplot <- leaflet(data=country_mpa2) %>% 
  addProviderTiles(providers$Stamen.Watercolor, options = providerTileOptions(noWrap = TRUE)) %>% 
  addPolygons(
    popup = ~sprintf("%s, %s", WDPAID, records),
    fillColor = ~pal(records),
      weight = 2,
      opacity = 1,
      color = "white",
      dashArray = "3",
      fillOpacity = 0.7) %>% 
  addLegend(pal = pal, values = ~records, opacity = 0.7, title = NULL,
  position = "bottomright")
leafletplot 
```





